Flowchart of Current Algorithm

### 2009-04-29 docs/flowchart/flowchart.tex
typo
changed some notation and added details about how to fix the gamma1bar problem in advect base spherical

\begin{color}{red}
Proposal: We need to pass in two different $\gammabar$ objects, $\gammabar^{\inp,n}$ and $\gammabar^{\inp,n+\myhalf}$.  The pressure update will use $\gammabar^{\inp,n+\myhalf}$.  To compute $\psi^{\outp,n}$, we use $\gammabar^{\inp,n}$.  To compute $\psi^{\outp,n+\myhalf}$, we use $\gammabar^{\inp,n+\myhalf}$.  In the first half of the algorithm, we pass in $\gammabar^n$ for both arguments.  For the second half of the algorithm, we pass in $\gammabar^n$ and $(\gammabar^n+\gammabar^{n+1,\star})/2$.  I don't think we ever need $\gammabar^{(1)}$, as the only place it's currently used is as an input to the first call to {\bf Advect Base Pressure Spherical}.
\end{color}

First use a predictor-corrector formulation for the pressure update 
in order to improve the overall accuracy.  Beginning with equation~(\ref{eq:w01}),
we can write
\begin{equation}
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} = 
 \Sbar - \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) - \frac{1}{\gammabar p_0} w_0 \frac{\partial p_0}{\partial r} \enskip .
\end{equation}
We time-center the left hand side:
\begin{equation}
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \approx
   \frac{2}{\gammabar^{\inp} (p_0^{\inp} + p_0^{\star})} \frac{p_0^{\star} - p_0^{\inp}}{\dt}\enskip.
\end{equation}
We can then solve for the provisional updated pressure, $p_0^{\star}$, giving:
\begin{equation}
p_0^{\star} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} f\right)}{\left(1 - \gammabar^{\inp} f\right)}\enskip,
\end{equation}
with:
\begin{equation}
f = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
                                             \frac{1}{\gammabar^{\inp} p_0^{\inp}} w_0^{\inp} \left ( \frac{\partial p_0^\inp}{\partial r} \right ) \right ] \label{eq:frhs} \enskip ,
\end{equation}
where the $(\partial p_0^\inp / \partial r)$ is computed by
interpolating $p_0^\inp$ from cell-centers to edges and then
differencing.  Now we begin the corrector step.  Using $p_0^\star$, we can compute a more 
accurate approximation to equation~(\ref{eq:frhs}),
\begin{equation}
\tilde{f} = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
                                             \frac{1}{\gammabar^{\inp} \tilde{p_0}} w_0^{\inp} \left ( \frac{\partial \tilde{p_0}}{\partial r} \right ) \right ] \enskip ,
\end{equation}
where $\tilde{p} = (p_0^\inp + p_0^\star)/2$, and the difference is again computed
by first putting $\tilde{p}$ on edges.

This allows us to complete the corrector for the pressure update:
\begin{equation}
p_0^{\outp} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} \tilde{f}\right)}{\left(1 - \gammabar^{\inp} \tilde{f}\right)}\enskip .
\end{equation}
We note that since we do not time-center $\gammabar$, this is not
strictly second-order accurate.  However, as $\gammabar$ varies slowly
in time, we expect the effect to be small.  This is something that we 
defer until simulations indicate it is warranted.

Next, we compute $\psi^{\outp,n}$ and $\psi^{\outp,n+\myhalf}$.  Rewriting Equation (\ref{eq:w01}), we can define
\begin{eqnarray}
\psi \equiv \frac{D_0 p_0}{Dt} &=& \gammabar p_0 \left [ \overline{S} - 
       \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) \right] \enskip .
\end{eqnarray}
We discretize this locally as:
\begin{equation}
\psi_j^{\outp,n} = 
\left(\gammabar^{\inp}p_0^{\inp}\right)_j
\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
          (r^2 w_0^{\outp})_{j-\myhalf} \right ] \right \} \enskip ,
\end{equation}
\begin{equation}
\psi_j^{\outp,n+\myhalf} = 
\left[\gammabar^{\inp} \frac{\left( p_0^{\inp} + p_0^{\outp} \right)}{2}\right]_j
\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
          (r^2 w_0^{\outp})_{j-\myhalf} \right ] \right \} \enskip .
\end{equation}
\end{description}

@@

\item {\bf Compute Gravity}$[\rhozero^{\inp}] \rightarrow [g^{\outp}]$\\
is the process by which we compute gravity at both radial cell-centers and edges.
\begin{description}
\item For plane-parallel, we hold the gravity constant in space and time.
\item For spherical, we define the enclosed mass at edges, and then cell-centers:
\begin{equation}
m_{j-\myhalf} = \sum_{k=1}^j \frac{4}{3}\pi\left(r_{k-\myhalf}^3 - r_{k-\sfrac{3}{2}}^3\right)\rho_{0,k-1}^{\inp}; ~~~~~ m_j = m_{j-\myhalf} + \frac{4}{3}\pi \left(r_k^3 - r_{k-\myhalf}^3\right)\rho_{0,k}^{\inp} \enskip .
\end{equation}
Then, we define the gravity at both edges and cell-centers:
\begin{equation}
g_{j-\myhalf} = \frac{Gm_{j-\myhalf}}{r_{j-\myhalf}^2}; ~~~~~ g_j = \frac{Gm_j}{r_j^2}\enskip .
\end{equation}
\end{description}

@@

\item Update $p_0$ and compute $\psi$.
\begin{description}
\item For plane-parallel,\\
{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph,\star}, \etarho^{n+\myhalf,\star}]
\rightarrow [p_0^{n+1,\star},\psi^{\nph,\star}]$.
\item For spherical,\\
{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph,\star},\gammabar^{(1)},
\overline{S^{\nph,\star}}] \rightarrow [p_0^{n+1,\star},\psi^{n,\star},\psi^{\nph,\star}]$.
\end{description}

@@

\item Update $p_0$ and compute $\psi$.
\begin{description}
\item For plane-parallel,\\
{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph}, \etarho^{n+\myhalf}]$ 
$\rightarrow [p_0^{n+1},\psi^{\nph}]$.
\item For spherical,\\
{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{n+1,\star},\overline{S^{\nph}}] \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.
\end{description}

>>>>

\begin{color}{red}
%Proposal: We need to pass in two different $\gammabar$ objects, $\gammabar^{\inp,n}$ and $\gammabar^{\inp,n+\myhalf}$.  The pressure update will use $\gammabar^{\inp,n+\myhalf}$.  To compute $\psi^{\outp,n}$, we use $\gammabar^{\inp,n}$.  To compute $\psi^{\outp,n+\myhalf}$, we use $\gammabar^{\inp,n+\myhalf}$.  In the first half of the algorithm, we pass in $\gammabar^n$ for both arguments.  For the second half of the algorithm, we pass in $\gammabar^n$ and $(\gammabar^n+\gammabar^{n+1,\star})/2$.  I don't think we ever need $\gammabar^{(1)}$, as the only place it's currently used is as an input to the first call to {\bf Advect Base Pressure Spherical}.
\end{color}

First use a predictor-corrector formulation for the pressure update 
in order to improve the overall accuracy.  Beginning with equation~(\ref{eq:w01}),
we can write
\begin{equation}
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} = 
 \Sbar - \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) - \frac{1}{\gammabar p_0} w_0 \frac{\partial p_0}{\partial r} \enskip .
\end{equation}
We time-center the left hand side:
\begin{equation}
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \approx
   \frac{2}{\gammabar^{\inp} (p_0^{\inp} + p_0^{\star})} \frac{p_0^{\star} - p_0^{\inp}}{\dt}\enskip.
\end{equation}
We can then solve for the provisional updated pressure, $p_0^{\star}$, giving:
\begin{equation}
p_0^{\star} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} f\right)}{\left(1 - \gammabar^{\inp} f\right)}\enskip,
\end{equation}
with:
\begin{equation}
f = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
                                             \frac{1}{\gammabar^{\inp} p_0^{\inp}} w_0^{\inp} \left ( \frac{\partial p_0^\inp}{\partial r} \right ) \right ] \label{eq:frhs} \enskip ,
\end{equation}
where the $(\partial p_0^\inp / \partial r)$ is computed by
interpolating $p_0^\inp$ from cell-centers to edges and then
differencing.  Now we begin the corrector step.
\begin{color}{red} We need to compute 
$\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{(2)},p_0^\star,X_k^{(2)})]$ 
and $\gammabar^\dagger = (\gammabar^\inp + \gammabar^\star)/2$, and then use 
$\gammabar^\dagger$ in the $f^\dagger$ equation. \end{color} 
Using $p_0^\star$, we can compute a more 
accurate approximation to equation~(\ref{eq:frhs}),
\begin{equation}
f^\dagger = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
                                             \frac{1}{\gammabar^{\inp} p_0^\dagger} w_0^{\inp} \left ( \frac{\partial p_0^\dagger}{\partial r} \right ) \right ] \enskip ,
\end{equation}
where $p_0^\dagger = (p_0^\inp + p_0^\star)/2$, and the difference is again computed
by first putting $p_0^\dagger$ on edges.

This allows us to complete the corrector for the pressure update:
\begin{equation}
p_0^{\outp} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} \tilde{f}\right)}{\left(1 - \gammabar^{\inp} \tilde{f}\right)}\enskip .
\end{equation}
We note that since we do not time-center $\gammabar$, this is not
strictly second-order accurate.  However, as $\gammabar$ varies slowly
in time, we expect the effect to be small.  This is something that we 
defer until simulations indicate it is warranted.\\

Next, we compute $\psi^{\outp,n}$ and $\psi^{\outp,n+\myhalf}$.  Rewriting Equation (\ref{eq:w01}), we can define
\begin{eqnarray}
\psi \equiv \frac{D_0 p_0}{Dt} &=& \gammabar p_0 \left [ \overline{S} - 
       \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) \right] \enskip .
\end{eqnarray}
We discretize this locally as:
\begin{equation}
\psi_j^{\outp,n} = 
\left(\gammabar^{\inp}p_0^{\inp}\right)_j
\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
          (r^2 w_0^{\inp})_{j-\myhalf} \right ] \right \} \enskip ,
\end{equation}
\begin{equation}
\psi_j^{\outp,n+\myhalf} = 
\left[\gammabar^{\inp} \frac{\left( p_0^{\inp} + p_0^{\outp} \right)}{2}\right]_j
\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
          (r^2 w_0^{\inp})_{j-\myhalf} \right ] \right \} \enskip .
\end{equation}
\end{description}

@@

\item {\bf Compute Gravity}$[\rhozero^{\inp}] \rightarrow [g^{\outp}]$\\
is the process by which we compute gravity at both radial cell-centers and edges.
\begin{description}
\item For plane-parallel, we hold the gravity constant in space and time.
\item For spherical, we define the enclosed mass at edges, and then cell-centers:
\begin{equation}
m_{j-\myhalf} = \sum_{k=1}^j \frac{4}{3}\pi\left(r_{k-\myhalf}^3 - r_{k-\sfrac{3}{2}}^3\right)\rho_{0,k-1}^{\inp}; ~~~~~ m_j = m_{j-\myhalf} + \frac{4}{3}\pi \left(r_j^3 - r_{j-\myhalf}^3\right)\rho_{0,j}^{\inp} \enskip .
\end{equation}
Then, we define the gravity at both edges and cell-centers:
\begin{equation}
g_{j-\myhalf} = \frac{Gm_{j-\myhalf}}{r_{j-\myhalf}^2}; ~~~~~ g_j = \frac{Gm_j}{r_j^2}\enskip .
\end{equation}
\end{description}

@@

\item Update $p_0$ and compute $\psi$.
\begin{description}
\item For plane-parallel,\\
{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph,\star}, \etarho^{n+\myhalf,\star}]
\rightarrow [p_0^{n+1,\star},\psi^{\nph,\star}]$.
\item For spherical,\\
{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph,\star},\gammabar^{(1)},
\overline{S^{\nph,\star}}] \rightarrow [p_0^{n+1,\star},\psi^{n,\star},\psi^{\nph,\star}]$.

@@

\item Update $p_0$ and compute $\psi$.
\begin{description}
\item For plane-parallel,\\
{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph}, \etarho^{n+\myhalf}]$ 
$\rightarrow [p_0^{n+1},\psi^{\nph}]$.
\item For spherical,\MarginPar{$\gammabar^{(1)}$ instead of $\gammabar^{n+1,\star}$?}\\
{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph},\gammabar^{n+1,\star},\overline{S^{\nph}}] \rightarrow [p_0^{n+1},\psi^n,\psi^{\nph}]$.
\end{description}
